Interactive Checks for Coordination Avoidance

### 2019-09-25 doc/vldb_slides/sections/bank_example.tex
Final edits.

\begin{frame}{Replicated Bank Account}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
    \end{tikzpicture}
  \end{center}

@@

    Let's say we have an object that we want to replicate. The object can be
    something complex like a database or a key-value store, but to keep things
    simple, let's say we want to replicate a single integer that represents my
    bank account balance. We decide to replicate the integer across two
    machines, replica A and replica B. \\[12pt]

    Users can issue transactions to our replicate object, and every
    transacation is either a deposit or a withdrawal. And, I have one critical
    invariant that I want to maintain at all times. And that invariant is that
    my bank account balance should never be negative. \\[12pt]
  }
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_1$}] (b1) at (0,  0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
      \pause

@@

  \note{%
    One way to replicate our object is by using coordination to implement
    strong consistency. For example, imagine a user issues a deposit of 10
    dollars to replica A, and simultaneously, another user issues a deposit of
    20 dollars to replica B. The two replicas can coordinate and agree on a
    single total order in which to execute the two transactions. Here, the
    replicas deposit 10 dollars and then deposit 20 dollars. Both transactions
    succeed and our invariant is maintained.
  }
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
      \pause

@@

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
      \pause

@@

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
      \pause

>>>>

\begin{frame}{Replicated Bank Account}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_2$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
    \end{tikzpicture}
  \end{center}

@@

    Let's say we have an object that we want to replicate. The object can be
    something complex like a database or a key-value store, but to keep things
    simple, let's say we want to replicate a single integer that represents my
    bank account balance. We decide to replicate the integer across two
    machines, $p_1$ and $p_2$. \\[12pt]

    Users can issue transactions to our replicated bank, and every
    transacation is either a deposit or a withdrawal. And, I have one critical
    invariant that I want to maintain at all times. And that invariant is that
    my bank account balance should never be negative. \\[12pt]
  }
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_2$}] (b1) at (0,  0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
      \pause

@@

  \note{%
    One way to replicate our object is by using coordination to implement
    strong consistency. For example, imagine a user issues a deposit of 10
    dollars to $p_1$, and simultaneously, another user issues a deposit of
    20 dollars to $p_2$. The two replicas can coordinate and agree on a
    single total order in which to execute the two transactions. Here, the
    replicas deposit 10 dollars and then deposit 20 dollars. Both transactions
    succeed and our invariant is maintained.
  }
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_2$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
      \pause

@@

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_2$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
      \pause

@@

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
      \node[bank, label=left:{$p_2$}] (b1) at (0, 0) {\$50};
      \draw[pline] (a1) to (9, \bankdistance);
      \draw[pline] (b1) to (9, 0);
      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
      \pause
